Continuation of the Swedish Civil Contingencies Agency (MSB) methodological support for \ac{isms}\\ MSB:s metodstöd

### 2017-04-07 msbframework/msbforts/msbforts.tex
Translated lecture
Signed-off-by: LENFRA <lennart.franked@miun.se>

\begin{frame}{What needs protecting?}
  \begin{itemize}
    \item What informational assets to we have, and how much are they worth
      protecting?
    \item Should lead to a structured list over
      \begin{itemize}
        \item what informational assets there are,,
        \item what requirements and expectations they have, and
        \item the worth of each asset.
      \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Formal methods?}
  \begin{itemize}
    \item There is some research done, looking into how to perform risk
      analysis.  
    \item They can, however be difficult to use.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Requirements for performing a gap analysis}
  The project leader of the analysis task must
  \begin{itemize}
    \item know the organisations need and requirements,
    \item know any existing governance documents for the security work,
    \item have a good knowledge and understanding of the norms in the standard.
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{itemize}
    \item There must exist a clear decision that a gap analysis should be
      performed,
    \item Otherwise there must be exist a clear mandate to be able to make this call.
    \item As a support, use MSB:s checklist for gap analysis~\cite{MSB2011gb}.
  \end{itemize}
\end{frame}

@@

\begin{frame}{ISO 27002}{Security Policy}
  \begin{itemize}
    \item In what direction do the organisation want to take the work within
      information security.
    \item How this should be achieved.
    \item Short explanation of terminology
    \item How the responsibility is divided.
    \item Legal and internal requirements should be taken into account.
  \end{itemize}
\end{frame}

\begin{frame}{ISO 27002}{Organisation surrounding information security}
  \begin{itemize}
    \item Should exist a frame work within the organisation that is controlling
      the information security.
    \item Governance documents should be decided.
    \item Clearly stated how responsibilities are divided.
    \item Contains six critical security measures.
  \end{itemize}
\end{frame}

@@

\begin{frame}{ISO 27002}{Compliance}
  \begin{itemize}
    \item Ensures that the organisation comply to the external and internal
      requirements that exist on the information security.
    \item Contains three critical security measures.
  \end{itemize}
\end{frame}

@@

\subsection{Processes}
\begin{frame}{Develop security processes}
  \begin{itemize}
    \item Collection of activities that manages a defined need.
    \item Integrate with existing processes.
    \item For example \ac{isms}.
    \item Examples: Information security classification, access control.
  \end{itemize}
\end{frame}
\begin{frame}{Developing processes}
  \begin{itemize}
    \item Wide competence.
    \item Representatives from the entire organisation.
    \item Makes it easier to coordinate with existing processes.
  \end{itemize}
\end{frame}

\subsection{Governance documents}
\begin{frame}{Develop governance documents}
  \begin{itemize}
    \item Start with one policy, in what direction the organisation wants to go.
    \item Identify existing documents: revise, revoke, introduce new.
    \item Adapt based on the current document structure.
  \end{itemize}
\end{frame}
\begin{frame}{Develop governance documents}
  Don't forget
  \begin{itemize}
    \item version control,
    \item document owner,
    \item decision date,
    \item who took the decision.
  \end{itemize}
  They should be written in such a way that everyone can understand them
\end{frame}

\section{Establish}
\begin{frame}{Establish}
  \begin{itemize}
    \item Security processes are formed and decisions for security measures
      exist.
    \item Can be done in project form, or just included into regular work load.
  \end{itemize}
\end{frame}

\subsection{Plan implementation}
\begin{frame}{Plan implementation}
  \begin{itemize}
    \item Should something be prioritized?
    \item Are some implementation measures dependent on others?
    \item What can be constructed, and what can be procured?
    \item Need a time plan, and ensure everyone is aware that the work will
      start.
  \end{itemize}
\end{frame}

\subsection{Construct and procure}
\begin{frame}{Construct and procure}
  \begin{description}
    \item[Construct] The organisation to in-house development.
    \item[Procure] The organisation procures solutions.
  \end{description}
  \begin{figure}
    \includegraphics[width=\textwidth]{infora.png}
    \caption{Process to construct and procure.}
  \end{figure}
\end{frame}
\begin{frame}{Defining requirements}
  \begin{itemize}
    \item Important to have clearly stated what requirements there are.
    \item IT-department are unaware of how much the information they are
      managing is worth protecting -- Need to be specified.
    \item Think about usability to ensure that the organisation do not
      circumvent the measures.
  \end{itemize}
\end{frame}

\subsection{Implement}
\begin{frame}{Implement}
  \begin{itemize}
    \item One coordinator.
    \item Multiple persons responsible for the practical.
    \item The work is carried out in groups: Include area managers that are
      affected by the security measures.
  \end{itemize}
\end{frame}
\begin{frame}{Groups responsible for implementing the security measures}
  Focus on:
  \begin{itemize}
    \item Representation of the entire organisation.
    \item Do the employees need training?
      They need to be able to use and understand the security measures.
    \item Who should manage the security measures?
      There is a need for a management plan.
    \item Ensure that the organisation will embrace the measures.
  \end{itemize}
\end{frame}
\begin{frame}{Changing the behaviour}
  \begin{itemize}
    \item Many security measures needs a change in the behaviour to work.
    \item Communicate and explain \emph{why} -- Make sure the explanation is
      targeted.
  \end{itemize}
\end{frame}

@@

\subsection{Review}
\begin{frame}{Review}
  \begin{itemize}
    \item A deeper analysis of the result from the monitoring.
    \item Part of the recurrent organisational follow-ups, for example internal
      revision.
    \item Use the information from the monitoring as base.
    \item Is the \ac{isms} working\@?
  \end{itemize}
\end{frame}
\begin{frame}{Audit requirement}
  \begin{itemize}
    \item ISO 27001: the top management must ensure this is done -- and should
      take part of the result.
    \item How?
      \begin{itemize}
        \item Use ISO 27001 \ac{isms}\@.
        \item Use ISO 27002 for the security measures.
      \end{itemize}
    \item Who?
      Independent auditor with the support from those who are responsible for
      the \ac{isms}.
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{What needs protecting?}
  \begin{itemize}
    \item What informational assets do we have, and how much are they worth
      protecting?
    \item Should lead to a structured list over
      \begin{itemize}
        \item what informational assets there are,,
        \item what requirements and expectations they have, and
        \item the worth of each asset.
      \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Formal methods?}
  \begin{itemize}
    \item There are a lot of research done regarding how to perform risk
      analysis.
    \item They can, however be difficult to use.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Requirements for performing a gap analysis}
  The project leader of the analysis task must
  \begin{itemize}
    \item know the organisations need and requirements,
    \item know any existing governance documents regarding the security work in
      the organisation,
    \item have a good knowledge and understanding of the norms in the standard.
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{itemize}
    \item There must exist a clear decision that a gap analysis should be
      performed,
    \item Otherwise there must exist a clear mandate to be able to make this call.
    \item As a support, use MSB:s checklist for gap analysis~\cite{MSB2011gb}.
  \end{itemize}
\end{frame}

@@

\begin{frame}{ISO 27002}{Security Policy}
  \begin{itemize}
    \item What is the directional intent for the organisation in regards to
      information security.
    \item How should this be achieved.
    \item Short explanation of terminology
    \item How will the responsibility be divided.
    \item Legal and internal requirements should be taken into account.
  \end{itemize}
\end{frame}

\begin{frame}{ISO 27002}{Organisation surrounding information security}
  \begin{itemize}
    \item Should exist a framework within the organisation that is controlling
      the information security.
    \item Decisions need to be taken regarding governance documents.
    \item Clearly stated how responsibilities are divided.
    \item Contains six critical security measures.
  \end{itemize}
\end{frame}

@@

\begin{frame}{ISO 27002}{Compliance}
  \begin{itemize}
    \item Ensures that the organisation comply to the external and internal
      requirements that exist in regards to information security.
    \item Contains three critical security measures.
  \end{itemize}
\end{frame}

@@

\subsection{Processes}
\begin{frame}{Develop security processes}
  \begin{itemize}
    \item Collection of activities that manages a defined need.
    \item Integrate with existing processes.
    \item For example \ac{isms}.
    \item Examples: Information security classification, access control.
  \end{itemize}
\end{frame}
\begin{frame}{Developing processes}
  \begin{itemize}
    \item Wide competence.
    \item Representatives from the entire (affected parts) organisation.
    \item Makes it easier to coordinate with existing processes.
  \end{itemize}
\end{frame}

\subsection{Governance documents}
\begin{frame}{Develop governance documents}
  \begin{itemize}
    \item Start with one policy, what is the directional intent of the organisation.
    \item Identify existing documents: revise, revoke, introduce new.
    \item Adapt based on the current document structure.
  \end{itemize}
\end{frame}
\begin{frame}{Develop governance documents}
  Don't forget
  \begin{itemize}
    \item version control,
    \item document owner,
    \item decision date,
    \item who took the decision.
  \end{itemize}
  They should be written in such a way that everyone can understand them
\end{frame}

\section{Implementation}
\begin{frame}{Implementation}
  \begin{itemize}
    \item At this stage security processes are formed and decisions have been
      made for taking security measures.
    \item Can be done in project form, or just incorporated into the regular work load.
  \end{itemize}
\end{frame}

\subsection{Plan the implementation}
\begin{frame}{Plan the implementation}
  \begin{itemize}
    \item Should something be prioritized?
    \item Do some implementation measures dependend on others?
    \item What can be constructed, and what can be procured?
    \item Need a time plan, and ensure everyone is aware that the work will
      start.
  \end{itemize}
\end{frame}

\subsection{Construct and procure}
\begin{frame}{Construct and procure}
  \begin{description}
    \item[Construct] The organisation to in-house development.
    \item[Procure] The organisation procures solutions.
  \end{description}
  \begin{figure}
    \includegraphics[width=\textwidth]{infora.png}
    \caption{Process to construct and procure.}
  \end{figure}
\end{frame}
\begin{frame}{Defining requirements}
  \begin{itemize}
    \item Important clearly have stated what requirements exist.
    \item IT-department are unaware of how much the information they are
      managing is worth protecting -- Need to be specified.
    \item Think about usability to ensure that the organisation do not
      circumvent the measures.
  \end{itemize}
\end{frame}

\subsection{Implement}
\begin{frame}{Implement}
  \begin{itemize}
    \item One coordinator.
    \item Multiple persons responsible for the practical.
    \item The work is carried out in groups: Include area managers that are
      affected by the security measures.
  \end{itemize}
\end{frame}
\begin{frame}{Groups responsible for implementing the security measures}
  Focus on:
  \begin{itemize}
    \item Representation of the entire organisation.
    \item Do the employees need training?
      They need to be able to use and understand the security measures.
    \item Who should manage the security measures?
      There is a need for a management plan.
    \item Ensuring that the organisation will embrace the measures.
  \end{itemize}
\end{frame}
\begin{frame}{Changing the behaviour}
  \begin{itemize}
    \item Many security measures needs a change in the behaviour to work.
    \item Communicate and explain \emph{why} -- Make sure the explanation is
      targeted.
  \end{itemize}
\end{frame}

@@

\subsection{Review}
\begin{frame}{Review}
  \begin{itemize}
    \item A deeper analysis of the result from the monitoring.
    \item Part of the recurring organisational follow-ups, for example internal
      revision.
    \item Use the information from the monitoring as base.
    \item Is the \ac{isms} working\@?
  \end{itemize}
\end{frame}
\begin{frame}{Audit requirement}
  \begin{itemize}
    \item ISO 27001: the top management must ensure this is done -- and should
      take part of the result.
    \item How?
      \begin{itemize}
        \item Use ISO 27001 \ac{isms}\@.
        \item Use ISO 27002 for the security measures.
      \end{itemize}
    \item Who?
      Independent auditor with the support from those who are responsible for
      the \ac{isms}.
  \end{itemize}
\end{frame}
